This PDF is a Git Repository\\Containing its Own \LaTeX\ Source\\and a Copy of Itself

### 2017-05-10 article.tex
Cleaned up the ISO standard code
Forgot to syntax color `return`!

One would
\emph{think} that git would use the same multi-byte length encoding that they
used for the uncompressed object length.  But no! This is what we have
to go off of from the git documentation:
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
n bytes with MSB set in all but the last one.
The offset is then the number constructed by
concatenating the lower 7 bit of each byte, and
for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))
to the result.
\end{verbatim}
\end{minipage}
\endgroup
\end{center}
Right. Some experimenting resulted in the following decoding logic
that appears to work:
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|        reference += (1 << (7 * (bytes_read - 1)))|}
\begin{minipage}{\wd9}
\footnotesize
{\color{blue}\verb|def|}\verb| decode_obj_ref|{\color{gray}\verb|(|}\verb|data|{\color{gray}\verb|):|}\\
\verb|    bytes_read = 0|\\
\verb|    reference = 0|\\
\verb|    |{\color{blue}\verb|for|}\verb| c |{\color{blue}\verb|in map|}{\color{gray}\verb|(|}\verb|ord, data|{\color{gray}\verb|):|}\\
\verb|        bytes_read += 1|\\
\verb|        reference <<= 7|\\
\verb|        reference += c & 0b01111111|\\
\verb|        |{\color{blue}\verb|if not|}\verb| |{\color{gray}\verb|(|}\verb|c & 0b10000000|{\color{gray}\verb|):|}\\
{\color{blue}\verb|            break|}\\
{\color{blue}\verb|    if|}\verb| bytes_read >= 2|{\color{gray}\verb|:|}\\
\verb|        reference += |{\color{gray}\verb|(|}\verb|1 << |{\color{gray}\verb|(|}\verb|7 * |{\color{gray}\verb|(|}\verb|bytes_read - 1|{\color{gray}\verb|)))|}\\
\verb|    return reference, bytes_read|\\
\end{minipage}
\endgroup
\end{center}

@@

Okay, I have to admit something: There is one more challenge. You see,
the latest PDF standard, ISO~32000-1:2008, says
\begin{center}
\begin{tikzpicture}
\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
The \emph{trailer} of a PDF file enables a conforming reader to
quickly find the cross-reference table and certain special
objects. Conforming readers should read a PDF file from its end. The
last line of the file shall contain only the end-of-file
marker, \texttt{\%\%EOF}.
\end{minipage}};
\node[anchor=north east] at (text.north west) {\LARGE``};
\node[anchor=south west] at (text.south east) {\LARGE''};
\end{tikzpicture}
\end{center}
Granted, we are producing a PDF that conforms to version~1.4 of the
specification, which doesn't appear to have that requirement.
However, at least as early as version 1.3, the specification did have
an implementation note that Acrobat requires the \texttt{\%\%EOF} to
be within the last 1024 bytes of the file.  Either way, that's not
guaranteed to be the case for us, especially since we are moving the
PDF to be at the beginning of the packfile.  There are almost
certainly going to be trailing bytes after the PDF's \texttt{\%\%EOF},
and if the git repository is large, those are likely to be more than
1024 bytes.

>>>>

One would
\emph{think} that git would use the same multi-byte length encoding that they
used for the uncompressed object length.  But no! This is what we have
to go off of from the git documentation:
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
n bytes with MSB set in all but the last one.
The offset is then the number constructed by
concatenating the lower 7 bit of each byte, and
for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))
to the result.
\end{verbatim}
\end{minipage}
\endgroup
\end{center}
Right. Some experimenting resulted in the following decoding logic
that appears to work:
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|        reference += (1 << (7 * (bytes_read - 1)))|}
\begin{minipage}{\wd9}
\footnotesize
{\color{blue}\verb|def|}\verb| decode_obj_ref|{\color{gray}\verb|(|}\verb|data|{\color{gray}\verb|):|}\\
\verb|    bytes_read = 0|\\
\verb|    reference = 0|\\
\verb|    |{\color{blue}\verb|for|}\verb| c |{\color{blue}\verb|in map|}{\color{gray}\verb|(|}\verb|ord, data|{\color{gray}\verb|):|}\\
\verb|        bytes_read += 1|\\
\verb|        reference <<= 7|\\
\verb|        reference += c & 0b01111111|\\
\verb|        |{\color{blue}\verb|if not|}\verb| |{\color{gray}\verb|(|}\verb|c & 0b10000000|{\color{gray}\verb|):|}\\
{\color{blue}\verb|            break|}\\
{\color{blue}\verb|    if|}\verb| bytes_read >= 2|{\color{gray}\verb|:|}\\
\verb|        reference += |{\color{gray}\verb|(|}\verb|1 << |{\color{gray}\verb|(|}\verb|7 * |{\color{gray}\verb|(|}\verb|bytes_read - 1|{\color{gray}\verb|)))|}\\
{\color{blue}\verb|    return|}\verb| reference, bytes_read|\\
\end{minipage}
\endgroup
\end{center}

@@

Okay, I have to admit something: There is one more challenge. You see,
the PDF standard~(ISO~32000-1) says
\begin{center}
\begin{tikzpicture}
\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
The \emph{trailer} of a PDF file enables a conforming reader to
quickly find the cross-reference table and certain special
objects. Conforming readers should read a PDF file from its end. The
last line of the file shall contain only the end-of-file
marker, \texttt{\%\%EOF}.
\end{minipage}};
\node[anchor=north east] at (text.north west) {\LARGE``};
\node[anchor=south west] at (text.south east) {\LARGE''};
\end{tikzpicture}
\end{center}
Granted, we are producing a PDF that conforms to version~1.4 of the
specification, which doesn't appear to have that requirement.
However, at least as early as version 1.3, the specification did have
an implementation note that Acrobat requires the \texttt{\%\%EOF} to
be within the last 1024 bytes of the file.  Either way, that's not
guaranteed to be the case for us, especially since we are moving the
PDF to be at the beginning of the packfile.  There are almost
certainly going to be trailing bytes after the PDF's \texttt{\%\%EOF},
and if the git repository is large, those are likely to be more than
1024 bytes.
